---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
set.seed(0)
```

# TransPhyloMulti

![GitHub](https://img.shields.io/github/license/DrJCarson/TransPhyloMulti)

# Introduction

TransPhyloMulti is an extension of [TransPhylo](https://github.com/xavierdidelot/TransPhylo)
which allows the use of multiple genomes per host. It also relaxes the assumption of a 
complete transmission bottleneck, and allows the within-host population to follow
a linear growth model.

For a more formal description of TransPhyloMulti, see the following paper:

Carson et al (2023). Inference of infectious disease transmission using multiple genomes per host. In preparation.

## Installation

You can install TransPhyloMulti from github with:

```{r gh-installation, eval = FALSE}
devtools::install_github("DrJCarson/TransPhyloMulti")
```

The package can then be loaded using:
```{r, eval = FALSE}
library(TransPhyloMulti)
```

## Basic usage

You will need as input a dated phylogeny built for example using [BEAST](https://www.beast2.org/), [BactDating](https://github.com/xavierdidelot/BactDating) or [treedater](https://github.com/emvolz/treedater). This dated phylogeny should be stored in the object `dated_tree` of class phylo from the ape package. Since this only includes relative rather than absolute dating, you also need to know the date when the last sample was taken, let's say on the 1st July 2022. 
You will also need a named vector `hosts` indicating the host
from which each of the leaves was sampled.
You can load and plot this data to make sure it looks correct:

```{r, eval = FALSE}
pt=ptreeFromPhyloM(dated_tree,lubridate::decimal_date(as.Date('2022/7/1')),hosts)
plot(pt)
```

You can then infer the transmission tree and 
associated parameters using for example:

```{r, eval = FALSE}
res=inferTTreeM(pt,w.shape=2,w.scale=2,obs.end=lubridate::decimal_date(as.Date('2023/1/1')))
plot(res)
```

Note that the parameters `w.shape` and `w.scale` specify a gamma distribution that needs to be appropriate to represent the generation time distribution of the pathogen under investigation. It is also important to input the date when sampling of cases ended, for example the 1st January 2023 in the code above.  

## More information and getting help

For more detailed examples of how to use TransPhyloMulti, see the vignettes [here](https://github.com/DrJCarson/TransPhyloMulti/tree/master/vignettes).
See also the help included in the package using the R command
`help(package='TransPhyloMulti')`. 

If you have any problem or question please create an issue [here](https://github.com/DrJCarson/TransPhyloMulti/issues) 
