---
title: "Exemplary Analysis of a Single Simulation"
author: "Jake Carson"
date: '`r Sys.Date()`'
output: pdf_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Exemplary Analysis of a Single Simulation}
  %\usepackage[utf8]{inputenc}
---

First we load the libraries we need using:
```{r}
library(ape)
library(TransPhylo)
library(TransPhyloMulti)
```

Set the simulation parameters.
```{r}
off.r <- 2
off.p <- 0.5
kappa <- 0.1
lambda <- 0.2
pi <- 0.8
sec.n <- 4
sec.t <- 0.25

w.shape <- 2
w.scale <- 1
ws.shape <- 2
ws.scale <- 1
outbreak.start <- 2000
obs.start <- 2000
grid.delta <- 1/365

host.lim <- 100
```

We set the random number generator so that results are perfectly reproducible.
```{r}
seed <- 10 * off.r + 1000 * pi + 10000 * kappa + 100000 * lambda + 3
set.seed(seed)
```

Simulate a dataset with the required number of hosts.
```{r}
repeat {

  sim_out <- sim_ongoing_lim_hosts(off.r = off.r,
                                   off.p = off.p,
                                   kappa = kappa,
                                   lambda = lambda,
                                   pi = pi,
                                   sec.n = sec.n,
                                   sec.t = sec.t,
                                   w.shape = w.shape,
                                   w.scale = w.scale,
                                   ws.shape = ws.shape,
                                   ws.scale = ws.scale,
                                   outbreak.start = outbreak.start,
                                   obs.start = obs.start,
                                   grid.delta = grid.delta,
                                   host.lim = host.lim)

  sim <- sim_out$ctree

  if (length(which(sim$ctree[, 2] == 0 & sim$ctree[, 3] == 0)) == host.lim * (1 + sec.n)) {
    
    break
    
  }
  
}
```

Extract observation cut-off time and dated phylogenetic tree.
```{r}
obs.end <- sim_out$obs.end
ptree <- extractPTreeM(sim)
```

Set MCMC chain length. To keep this vignette quick to compile we will do only 1e2 iterations but this should be replaced with 12000 to match the paper.
```{r}
mcmc.length = 100
```

Run MCMC for full dataset.
```{r}
sim5 <- sim
ptree5 <- extractPTreeM(sim5)
parms_out_5 <- inferTTreeM(ptree = ptree5, w.shape = w.shape, w.scale = w.scale, 
                           ws.shape = ws.shape, ws.scale = ws.scale, 
                           obs.end = obs.end, verbose = F,
                             mcmc.length = mcmc.length, mcmc.thinning = 1)
```

Functions to trim samples from phylogenetic tree.
```{r}
order_hosts_2 <- function(ctree) {
  unique_hosts <- unique(ctree[, 4])
  n_hosts <- length(unique_hosts) - 1
  ordered_hosts <- order(unique_hosts[1:n_hosts])
  ctree[which(ctree[, 4] > 0), 4] <- ordered_hosts[ctree[which(ctree[, 4] > 0), 4]]
  return(ctree)
}

drop_obs <- function(ctree, maxobs = 1) {
  nam <- ctree$nam
  ctree <- ctree$ctree
  remleaves <- rep(F, length(which(ctree[, 2] == 0 & ctree[, 3] == 0)))
  n_hosts <- max(ctree[, 4])
  for (h in 1:n_hosts) {
    r <- which(ctree[, 2] == 0 & ctree[, 3] == 0 & ctree[, 4] == h)
    if (length(r) > maxobs) {
      remleaves[r[order(ctree[r, 1], decreasing = T)[1:(length(r) - maxobs)]]] <- T
    }
  }
  wrl=which(remleaves)
  nam <- nam[remleaves == F]
  rem=c(remleaves,rep(F,nrow(ctree)-length(remleaves)))
  parents=rep(NA,nrow(ctree)+1)
  parents[ctree[,2]+1]=1:nrow(ctree)
  parents[ctree[,3]+1]=1:nrow(ctree)
  parents=parents[-1]
  for (j in wrl) {
    cur=j
    while (rem[cur]==T || ctree[cur,3]==0) {
      rem[cur]=T
      cur=parents[cur]
    }
    rem[cur]=T
  }
  w=which(!rem)
  map=rep(NA,nrow(ctree))
  map[w]=1:length(w)
  for (j in 1:length(map)) {
    cur=j
    while (all(is.na(map[cur]))) {cur=ctree[cur,2:3];cur=cur[!is.na(cur)&cur>0]
    if (length(cur)==0) break;}
    if (length(cur)>0) {tmp=map[cur];map[j]=tmp[!is.na(tmp)]}
  }
  ctree=ctree[!rem,]
  w=which(ctree[,2]>0);ctree[w,2]=map[ctree[w,2]]
  w=which(ctree[,3]>0);ctree[w,3]=map[ctree[w,3]]
  ctree <- order_hosts_2(ctree)
  l <- list(ctree = ctree, nam = nam)
  class(l) <- 'ctree'
  return(l)
}
```

Create phylogenetic trees with 4, 3, 2, 1 observations per host and repeat the analysis.
```{r}
sim4 <- drop_obs(sim, maxobs = 4)
ptree4 <- extractPTreeM(sim4)
  parms_out_4 <- inferTTreeM(ptree = ptree4, w.shape = w.shape, w.scale = w.scale, 
                             ws.shape = ws.shape, ws.scale = ws.scale, 
                             obs.end = obs.end, verbose = F, 
                             mcmc.length = mcmc.length, mcmc.thinning = 1)

sim3 <- drop_obs(sim, maxobs = 3)
ptree3 <- extractPTreeM(sim3)
parms_out_3 <- inferTTreeM(ptree = ptree3, w.shape = w.shape, w.scale = w.scale, 
                           ws.shape = ws.shape, ws.scale = ws.scale, 
                           obs.end = obs.end, verbose = F, 
                           mcmc.length = mcmc.length, mcmc.thinning = 1)

sim2 <- drop_obs(sim, maxobs = 2)
ptree2 <- extractPTreeM(sim2)
parms_out_2 <- inferTTreeM(ptree = ptree2, w.shape = w.shape, w.scale = w.scale, 
                           ws.shape = ws.shape, ws.scale = ws.scale, 
                           obs.end = obs.end, verbose = F,
                           mcmc.length = mcmc.length, mcmc.thinning = 1)

sim1 <- drop_obs(sim, maxobs = 1)
ptree1 <- extractPTreeM(sim1)
parms_out_1 <- inferTTreeM(ptree = ptree1, w.shape = w.shape, w.scale = w.scale, 
                           ws.shape = ws.shape, ws.scale = ws.scale, 
                           obs.end = obs.end, verbose = F,
                           mcmc.length = mcmc.length, mcmc.thinning = 1)

```

Set burn-in and indices for posterior samples.
```{r}
burn.in <- floor((1 / 6) * mcmc.length)
post.ind <- burn.in:mcmc.length
```

Mean and credible intervals for pi.
```{r}
pi1 <- sapply(parms_out_1, function(x) x[["pi"]])[post.ind]
pi2 <- sapply(parms_out_2, function(x) x[["pi"]])[post.ind]
pi3 <- sapply(parms_out_3, function(x) x[["pi"]])[post.ind]
pi4 <- sapply(parms_out_4, function(x) x[["pi"]])[post.ind]
pi5 <- sapply(parms_out_5, function(x) x[["pi"]])[post.ind]

pi_array <- array(numeric(15), dim = c(5, 3))
pi_array[1, ] <- c(mean(pi1), sort(pi1)[c(floor(0.025 * length(post.ind)) + 1, 
                                          floor(0.975 * length(post.ind)))])
pi_array[2, ] <- c(mean(pi2), sort(pi2)[c(floor(0.025 * length(post.ind)) + 1, 
                                          floor(0.975 * length(post.ind)))])
pi_array[3, ] <- c(mean(pi3), sort(pi3)[c(floor(0.025 * length(post.ind)) + 1, 
                                          floor(0.975 * length(post.ind)))])
pi_array[4, ] <- c(mean(pi4), sort(pi4)[c(floor(0.025 * length(post.ind)) + 1, 
                                          floor(0.975 * length(post.ind)))])
pi_array[5, ] <- c(mean(pi5), sort(pi5)[c(floor(0.025 * length(post.ind)) + 1, 
                                          floor(0.975 * length(post.ind)))])

print(pi_array)

```

Mean and credible intervals for off.r.
```{r}
offr1 <- sapply(parms_out_1, function(x) x[["off.r"]])[post.ind]
offr2 <- sapply(parms_out_2, function(x) x[["off.r"]])[post.ind]
offr3 <- sapply(parms_out_3, function(x) x[["off.r"]])[post.ind]
offr4 <- sapply(parms_out_4, function(x) x[["off.r"]])[post.ind]
offr5 <- sapply(parms_out_5, function(x) x[["off.r"]])[post.ind]

r_array <- array(numeric(15), dim = c(5, 3))
r_array[1, ] <- c(mean(offr1), sort(offr1)[c(floor(0.025 * length(post.ind)) + 1, 
                                             floor(0.975 * length(post.ind)))])
r_array[2, ] <- c(mean(offr2), sort(offr2)[c(floor(0.025 * length(post.ind)) + 1, 
                                             floor(0.975 * length(post.ind)))])
r_array[3, ] <- c(mean(offr3), sort(offr3)[c(floor(0.025 * length(post.ind)) + 1, 
                                             floor(0.975 * length(post.ind)))])
r_array[4, ] <- c(mean(offr4), sort(offr4)[c(floor(0.025 * length(post.ind)) + 1, 
                                             floor(0.975 * length(post.ind)))])
r_array[5, ] <- c(mean(offr5), sort(offr5)[c(floor(0.025 * length(post.ind)) + 1, 
                                             floor(0.975 * length(post.ind)))])

print(r_array)

```

Mean and credible intervals for kappa.
```{r}
kappa1 <- sapply(parms_out_1, function(x) x[["kappa"]])[post.ind]
kappa2 <- sapply(parms_out_2, function(x) x[["kappa"]])[post.ind]
kappa3 <- sapply(parms_out_3, function(x) x[["kappa"]])[post.ind]
kappa4 <- sapply(parms_out_4, function(x) x[["kappa"]])[post.ind]
kappa5 <- sapply(parms_out_5, function(x) x[["kappa"]])[post.ind]

k_array <- array(numeric(15), dim = c(5, 3))
k_array[1, ] <- c(mean(kappa1), sort(kappa1)[c(floor(0.025 * length(post.ind)) + 1, 
                                               floor(0.975 * length(post.ind)))])
k_array[2, ] <- c(mean(kappa2), sort(kappa2)[c(floor(0.025 * length(post.ind)) + 1, 
                                               floor(0.975 * length(post.ind)))])
k_array[3, ] <- c(mean(kappa3), sort(kappa3)[c(floor(0.025 * length(post.ind)) + 1, 
                                               floor(0.975 * length(post.ind)))])
k_array[4, ] <- c(mean(kappa4), sort(kappa4)[c(floor(0.025 * length(post.ind)) + 1,
                                               floor(0.975 * length(post.ind)))])
k_array[5, ] <- c(mean(kappa5), sort(kappa5)[c(floor(0.025 * length(post.ind)) + 1, 
                                               floor(0.975 * length(post.ind)))])

print(k_array)

```

Mean and credible intervals for lambda.
```{r}
lambda1 <- sapply(parms_out_1, function(x) x[["lambda"]])[post.ind]
lambda2 <- sapply(parms_out_2, function(x) x[["lambda"]])[post.ind]
lambda3 <- sapply(parms_out_3, function(x) x[["lambda"]])[post.ind]
lambda4 <- sapply(parms_out_4, function(x) x[["lambda"]])[post.ind]
lambda5 <- sapply(parms_out_5, function(x) x[["lambda"]])[post.ind]

l_array <- array(numeric(15), dim = c(5, 3))
l_array[1, ] <- c(mean(lambda1), sort(lambda1)[c(floor(0.025 * length(post.ind)) + 1, 
                                                 floor(0.975 * length(post.ind)))])
l_array[2, ] <- c(mean(lambda2), sort(lambda2)[c(floor(0.025 * length(post.ind)) + 1, 
                                                 floor(0.975 * length(post.ind)))])
l_array[3, ] <- c(mean(lambda3), sort(lambda3)[c(floor(0.025 * length(post.ind)) + 1, 
                                                 floor(0.975 * length(post.ind)))])
l_array[4, ] <- c(mean(lambda4), sort(lambda4)[c(floor(0.025 * length(post.ind)) + 1, 
                                                 floor(0.975 * length(post.ind)))])
l_array[5, ] <- c(mean(lambda5), sort(lambda5)[c(floor(0.025 * length(post.ind)) + 1, 
                                                 floor(0.975 * length(post.ind)))])

print(l_array)

```

Calculate posterior probabilities for who infected who with one and five observations per host.
```{r}
infm_1 <- matrix(0, nrow = 100, ncol = 100)
for (m in post.ind) {
  ctree <- parms_out_1[[m]]$ctree
  ttree <- extractTTreeM(ctree)$ttree
  for (h2 in 1:100) {
    h1 <- ttree[h2, 3]
    if (h1 %in% 1:100) {
      infm_1[h1, h2] <- infm_1[h1, h2] + 1
    }
  }
}
infm_1 <- infm_1 / length(post.ind)

infm_5 <- matrix(0, nrow = 100, ncol = 100)
for (m in post.ind) {
  ctree <- parms_out_5[[m]]$ctree
  ttree <- extractTTreeM(ctree)$ttree
  for (h2 in 1:100) {
    h1 <- ttree[h2, 3]
    if (h1 %in% 1:100) {
      infm_5[h1, h2] <- infm_5[h1, h2] + 1
    }
  }
}
infm_5 <- infm_5 / length(post.ind)

```

Determine who infected who in the simulation.
```{r}
hh_infected <- which(extractTTreeM(sim)$ttree[1:dim(infm_1)[1], 3] <= dim(infm_1)[1] & 
                       extractTTreeM(sim)$ttree[1:dim(infm_1)[1], 3] != 0)
hh_infector <- extractTTreeM(sim)$ttree[hh_infected, 3]
```

Plot posterior probabilities of who infected who using full dataset.
```{r}
lcols <- colorRampPalette(c("white", "red"))(20)
lattice::levelplot(infm_5, at = seq(0, 1, by = 0.05), col.regions = lcols,
                   xlab = "Infector", ylab = "Infected")
lattice::trellis.focus("panel", 1, 1, highlight = F)
lattice::lrect(xleft = hh_infector - 0.5, xright = hh_infector + 0.5, 
      ybottom = hh_infected - 0.5, ytop = hh_infected + 0.5,
      lwd = 2)
lattice::lrect(xleft = hh_infected - 0.5, xright = hh_infected + 0.5, 
      ybottom = hh_infector - 0.5, ytop = hh_infector + 0.5,
      lwd = 2, border = "gray70")
lattice::trellis.unfocus()
```

Show difference in posterior probabilities of who infected who from using one observation per host.
```{r}
lcols <- c(colorRampPalette(c("blue", "white"))(10), colorRampPalette(c("white", "red"))(10))
lattice::levelplot(infm_5 - infm_1, at = seq(-1, 1, by = 0.1), col.regions = lcols,
                   xlab = "Infector", ylab = "Infected")
lattice::trellis.focus("panel", 1, 1, highlight = F)
lattice::lrect(xleft = hh_infector - 0.5, xright = hh_infector + 0.5, 
      ybottom = hh_infected - 0.5, ytop = hh_infected + 0.5,
      lwd = 2)
lattice::lrect(xleft = hh_infected - 0.5, xright = hh_infected + 0.5, 
      ybottom = hh_infector - 0.5, ytop = hh_infector + 0.5,
      lwd = 2, border = "gray70")
lattice::trellis.unfocus()

```


Set thresholds for identifying transmission links.
```{r}
sen_threshold <- 0.5
spec_threshold <- 0.5
```

Calculate senstivities and specificies of identifying transmission links when using one observation per host.
```{r}

hh_probs_fwd_1 <- numeric(length(hh_infected))
hh_probs_bwd_1 <- numeric(length(hh_infected))
for (i in 1:length(hh_infected)) {
  hh_probs_fwd_1[i] <- infm_1[hh_infector[i], hh_infected[i]]
  hh_probs_bwd_1[i] <- infm_1[hh_infected[i], hh_infector[i]]
}
hh_probs_tot_1 <- hh_probs_fwd_1 + hh_probs_bwd_1
sen_tot_1 <- length(which(hh_probs_tot_1 >= sen_threshold)) / length(hh_probs_tot_1)
sen_fwd_1 <- length(which(hh_probs_fwd_1 >= sen_threshold)) / length(hh_probs_fwd_1)

fal_pos_fwd_1 <- 0
count_fwd_1 <- 0
fal_pos_tot_1 <- 0
count_tot_1 <- 0
nh_probs_tot_1 <- c()
nh_probs_fwd_1 <- c()
for (i in 1:100) {
  fwd_idx <- which(hh_infected == i)
  bwd_idx <- which(hh_infector == i)
  if (length(fwd_idx) > 0 & length(bwd_idx) > 0) {
    fal_pos_fwd_1 <- fal_pos_fwd_1 + 
      length(which(infm_1[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_fwd_1 <- count_fwd_1 + length(infm_1[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_fwd_1 <- c(nh_probs_fwd_1, infm_1[-c(i, hh_infector[fwd_idx]) , i])
    fal_pos_tot_1 <- fal_pos_tot_1 + 
      length(which(infm_1[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i] >= spec_threshold))
    count_tot_1 <- count_tot_1 + 
      length(infm_1[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i])
    nh_probs_tot_1 <- c(nh_probs_tot_1, infm_1[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i])
  } else if (length(bwd_idx) > 0) {
    fal_pos_fwd_1 <- fal_pos_fwd_1 + length(which(infm_1[-i , i] >= spec_threshold))
    count_fwd_1 <- count_fwd_1 + length(infm_1[-i , i])
    nh_probs_fwd_1 <- c(nh_probs_fwd_1, infm_1[-i , i])
    fal_pos_tot_1 <- fal_pos_tot_1 + 
      length(which(infm_1[-c(i, hh_infected[bwd_idx]) , i] >= spec_threshold))
    count_tot_1 <- count_tot_1 + length(infm_1[-c(i, hh_infected[bwd_idx]) , i])
    nh_probs_tot_1 <- c(nh_probs_tot_1, infm_1[-c(i, hh_infected[bwd_idx]) , i])
  } else if (length(fwd_idx) > 0) {
    fal_pos_fwd_1 <- fal_pos_fwd_1 + 
      length(which(infm_1[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_fwd_1 <- count_fwd_1 + length(infm_1[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_fwd_1 <- c(nh_probs_fwd_1, infm_1[-c(i, hh_infector[fwd_idx]) , i])
    fal_pos_tot_1 <- fal_pos_tot_1 + 
      length(which(infm_1[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_tot_1 <- count_tot_1 + length(infm_1[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_tot_1 <- c(nh_probs_tot_1, infm_1[-c(i, hh_infector[fwd_idx]) , i])
  } else {
    fal_pos_fwd_1 <- fal_pos_fwd_1 + length(which(infm_1[-i , i] >= spec_threshold))
    count_fwd_1 <- count_fwd_1 + length(infm_1[-i , i])
    nh_probs_fwd_1 <- c(nh_probs_fwd_1, infm_1[-i , i])
    fal_pos_tot_1 <- fal_pos_tot_1 + length(which(infm_1[-i , i] >= spec_threshold))
    count_tot_1 <- count_tot_1 + length(infm_1[-i , i])
    nh_probs_tot_1 <- c(nh_probs_tot_1, infm_1[-i , i])
  }
}
spec_fwd_1 <- 1 - fal_pos_fwd_1 / count_fwd_1
spec_tot_1 <- 1 - fal_pos_tot_1 / count_tot_1

print(paste("Bidirectional sensitivity: ", sen_tot_1, ".", sep = ""))
print(paste("Bidirectional specificity: ", spec_tot_1, ".", sep = ""))
print(paste("Directional sensitivity: ", sen_fwd_1, ".", sep = ""))
print(paste("Directional specificity: ", spec_fwd_1, ".", sep = ""))

```

Calculate senstivities and specificies of identifying transmission links when using five observations per host.
```{r}

hh_probs_fwd_5 <- numeric(length(hh_infected))
hh_probs_bwd_5 <- numeric(length(hh_infected))
for (i in 1:length(hh_infected)) {
  hh_probs_fwd_5[i] <- infm_5[hh_infector[i], hh_infected[i]]
  hh_probs_bwd_5[i] <- infm_5[hh_infected[i], hh_infector[i]]
}
hh_probs_tot_5 <- hh_probs_fwd_5 + hh_probs_bwd_5
sen_tot_5 <- length(which(hh_probs_tot_5 >= sen_threshold)) / length(hh_probs_tot_5)
sen_fwd_5 <- length(which(hh_probs_fwd_5 >= sen_threshold)) / length(hh_probs_fwd_5)

fal_pos_fwd_5 <- 0
count_fwd_5 <- 0
fal_pos_tot_5 <- 0
count_tot_5 <- 0
nh_probs_tot_5 <- c()
nh_probs_fwd_5 <- c()
for (i in 1:100) {
  fwd_idx <- which(hh_infected == i)
  bwd_idx <- which(hh_infector == i)
  if (length(fwd_idx) > 0 & length(bwd_idx) > 0) {
    fal_pos_fwd_5 <- fal_pos_fwd_5 + 
      length(which(infm_5[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_fwd_5 <- count_fwd_5 + length(infm_5[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_fwd_5 <- c(nh_probs_fwd_5, infm_5[-c(i, hh_infector[fwd_idx]) , i])
    fal_pos_tot_5 <- fal_pos_tot_5 + 
      length(which(infm_5[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i] >= spec_threshold))
    count_tot_5 <- count_tot_5 + 
      length(infm_5[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i])
    nh_probs_tot_5 <- c(nh_probs_tot_5, infm_5[-c(i, hh_infector[fwd_idx], hh_infected[bwd_idx]) , i])
  } else if (length(bwd_idx) > 0) {
    fal_pos_fwd_5 <- fal_pos_fwd_5 + length(which(infm_5[-i , i] >= spec_threshold))
    count_fwd_5 <- count_fwd_5 + length(infm_5[-i , i])
    nh_probs_fwd_5 <- c(nh_probs_fwd_5, infm_5[-i , i])
    fal_pos_tot_5 <- fal_pos_tot_5 + 
      length(which(infm_5[-c(i, hh_infected[bwd_idx]) , i] >= spec_threshold))
    count_tot_5 <- count_tot_5 + length(infm_5[-c(i, hh_infected[bwd_idx]) , i])
    nh_probs_tot_5 <- c(nh_probs_tot_5, infm_5[-c(i, hh_infected[bwd_idx]) , i])
  } else if (length(fwd_idx) > 0) {
    fal_pos_fwd_5 <- fal_pos_fwd_5 + 
      length(which(infm_5[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_fwd_5 <- count_fwd_5 + length(infm_5[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_fwd_5 <- c(nh_probs_fwd_5, infm_5[-c(i, hh_infector[fwd_idx]) , i])
    fal_pos_tot_5 <- fal_pos_tot_5 + 
      length(which(infm_5[-c(i, hh_infector[fwd_idx]) , i] >= spec_threshold))
    count_tot_5 <- count_tot_5 + length(infm_5[-c(i, hh_infector[fwd_idx]) , i])
    nh_probs_tot_5 <- c(nh_probs_tot_5, infm_5[-c(i, hh_infector[fwd_idx]) , i])
  } else {
    fal_pos_fwd_5 <- fal_pos_fwd_5 + length(which(infm_5[-i , i] >= spec_threshold))
    count_fwd_5 <- count_fwd_5 + length(infm_5[-i , i])
    nh_probs_fwd_5 <- c(nh_probs_fwd_5, infm_5[-i , i])
    fal_pos_tot_5 <- fal_pos_tot_5 + length(which(infm_5[-i , i] >= spec_threshold))
    count_tot_5 <- count_tot_5 + length(infm_5[-i , i])
    nh_probs_tot_5 <- c(nh_probs_tot_5, infm_5[-i , i])
  }
}

spec_fwd_5 <- 1 - fal_pos_fwd_5 / count_fwd_5
spec_tot_5 <- 1 - fal_pos_tot_5 / count_tot_5

print(paste("Bidirectional sensitivity: ", sen_tot_5, ".", sep = ""))
print(paste("Bidirectional specificity: ", spec_tot_5, ".", sep = ""))
print(paste("Directional sensitivity: ", sen_fwd_5, ".", sep = ""))
print(paste("Directional specificity: ", spec_fwd_5, ".", sep = ""))

```

Plot posterior link probabilities for one and five obserations per host for comparison.
```{r}

seny1 <- numeric(10)
seny5 <- numeric(10)
spec1 <- numeric(10)
spec5 <- numeric(10)
for (i in 1:10) {
  seny1[i] <- length(which(hh_probs_tot_1 >= (i - 1) * 0.1 & hh_probs_tot_1 < i * 0.1)) / 
    length(hh_probs_tot_1)
  seny5[i] <- length(which(hh_probs_tot_5 >= (i - 1) * 0.1 & hh_probs_tot_5 < i * 0.1)) / 
    length(hh_probs_tot_5)
  spec1[i] <- length(which(nh_probs_tot_1 >= (i - 1) * 0.1 & nh_probs_tot_1 < i * 0.1)) / 
    length(nh_probs_tot_1)
  spec5[i] <- length(which(nh_probs_tot_5 >= (i - 1) * 0.1 & nh_probs_tot_5 < i * 0.1)) / 
    length(nh_probs_tot_5)
}

par(mfrow = c(2, 2), lwd = 2, bty = "n", cex.axis = 1.2, cex.lab = 1.4, cex.main = 1.6)

plot(c(0, 1), c(0, 0.25), col = "white", xlab = "Bidirectional link probability", 
     ylab = "Proportion", main = "One observation per host")
x1 <- -0.002
y1 <- seny1[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "red2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- seny1[i]
  lines(c(x1, x2), c(y1, y2), col = "red2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "red2")
    break
  }
}
x1 <- 0.002
y1 <- spec1[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- spec1[i]
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "blue2")
    break
  }
}

plot(c(0, 1), c(0, 0.25), col = "white", xlab = "Bidirectional link probability", 
     ylab = "Proportion", main = "Five observations per host")
x1 <- -0.002
y1 <- seny5[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "red2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- seny5[i]
  lines(c(x1, x2), c(y1, y2), col = "red2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "red2")
    break
  }
}
x1 <- 0.002
y1 <- spec5[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- spec5[i]
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "blue2")
    break
  }
}

seny1 <- numeric(10)
seny5 <- numeric(10)
spec1 <- numeric(10)
spec5 <- numeric(10)
for (i in 1:10) {
  seny1[i] <- length(which(hh_probs_fwd_1 >= (i - 1) * 0.1 & hh_probs_fwd_1 < i * 0.1)) / 
    length(hh_probs_fwd_1)
  seny5[i] <- length(which(hh_probs_fwd_5 >= (i - 1) * 0.1 & hh_probs_fwd_5 < i * 0.1)) / 
    length(hh_probs_fwd_5)
  spec1[i] <- length(which(nh_probs_fwd_1 >= (i - 1) * 0.1 & nh_probs_fwd_1 < i * 0.1)) / 
    length(nh_probs_fwd_1)
  spec5[i] <- length(which(nh_probs_fwd_5 >= (i - 1) * 0.1 & nh_probs_fwd_5 < i * 0.1)) / 
    length(nh_probs_fwd_5)
}

plot(c(0, 1), c(0, 0.25), col = "white", xlab = "Directional link probability", 
     ylab = "Proportion", main = "")
x1 <- -0.002
y1 <- seny1[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "red2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- seny1[i]
  lines(c(x1, x2), c(y1, y2), col = "red2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "red2")
    break
  }
}
x1 <- 0.002
y1 <- spec1[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- spec1[i]
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "blue2")
    break
  }
}

plot(c(0, 1), c(0, 0.25), col = "white", xlab = "Directional link probability", 
     ylab = "Proportion", main = "")
x1 <- -0.002
y1 <- seny5[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "red2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- seny5[i]
  lines(c(x1, x2), c(y1, y2), col = "red2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "red2")
    break
  }
}
x1 <- 0.002
y1 <- spec5[1]
i <- 1
repeat {
  x2 <- x1 + 0.1
  y2 <- y1
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  x1 <- x2
  y1 <- y2
  i <- i + 1
  y2 <- spec5[i]
  lines(c(x1, x2), c(y1, y2), col = "blue2")
  y1 <- y2
  if (i == 10) {
    x2 <- x1 + 0.1
    y2 <- y1
    lines(c(x1, x2), c(y1, y2), col = "blue2")
    break
  }
}


```









