---
title: "Testing TransPhyloMulti"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output: pdf_document
---
```{r echo=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```

If you want to reproduce exactly the same results as the ones shown in this vignette, you need to set the seed of your random number generator to zero:
```{r}
library(TransPhylo)
library(TransPhyloMulti)
set.seed(0)
```

The offspring distribution is negative binomial with mean equal to the basic reproduction number $R=2$. Both the generation time and the sampling time are Gamma distributed with parameters (10,0.1) which has a mean of 1 year. The density of sampling is $\pi=0.25$. The following commands specify these parameters:
```{r}
off.r=2
w.shape=10
w.scale=0.1
pi=0.25
```

We simulate an outbreak that starts in 2005 and which and is observed up to 2010:

```{r}
dateT=2013
simu <- sim_ongoing(pi=pi,off.r=off.r,w.shape=w.shape,
                         w.scale=w.scale,dateStartOutbreak=2005,dateT=dateT,add.prob = 0.6,lm_const = 0.5,lm_rate = 0.5)
```

This simulation contains both the transmission tree between infected hosts and the within-host phylogenetic tree of each host. This can be visualised as a colored phylogenetic tree, where each host is represented by a unique color:

```{r}
plot(simu)
```

The transmission tree can be extracted and plotted separately from the phylogeny:

```{r}
ttree<-extractTTreeM(simu)
plot(ttree)
```

A more detailed plot can be displayed as follows:

```{r}
plot(ttree,type='detailed',w.shape,w.scale)
```

The phylogenetic tree can be extracted and plotted separately from the transmission tree:

```{r}
ptree<-extractPTreeM(simu)
plot(ptree)
```

The extracted phylogenetic tree can also be converted into a phylo object from the ape package:

```{r}
library(ape)
p<-phyloFromPTree(ptree)
plot(p)
axisPhylo(backward = F)
```


#```{r}
#write.tree(p,'tree.nwk')
#p=read.tree('tree.nwk')
#ptree=ptreeFromPhyloM(p,dateLastSample = dateLastSample(simu))
#```

Now let's try inference:

```{r,results='hide'}
res<-inferTTreeM(ptree,mcmcIterations=100,w.shape=w.shape,w.scale=w.scale,dateT=dateT,startPi=0.25,updatePi=T,startOff.r=2,updateOff.r=T,start_const=1,start_rate=1,updateconst = T,updaterate=T,startCTree = simu)
```

```{r}
plot(res)
```

